---
title: "isomiR annotation with STAR using whole genome"
author: "Lorena Pantano"
date: "11/29/2014"
output:
  knitrBootstrap::bootstrap_document:
    theme: readable
    highlight: zenburn
    theme.chooser: TRUE
    highlight.chooser: TRUE
  html_document:
    highlight: zenburn

---


```{r setup, echo=FALSE}
library(knitr)
library(rmarkdown)
library(knitrBootstrap)

opts_chunk$set(tidy=TRUE, highlight=T, figalign="center",
 fig.height=6, fig.width=6, message=F, error=F, warning=F, cache=T)
```


```{r load}
data<-read.table("sim.20.hsa.dat", header=T)
data_filtered<-read.table("sim.20.hsa.primary.dat", header=T)
```

# Overview

The idea of this report is to show how well STAR is decting
isomiRs using the genome as reference. I simulated a bunch of isomirs (~16900) that can have the folowing variation:

* starts at different position than the reference miRNA: t5
* ends at different position than the reference miRNA: t3
* have a mutation: muts
* have nt addition at the end: add

## Methods


`bash map_star.sh sim.20.hsa.fa`
`bamtools filter -isPrimaryAlignment True -in sim.20.hsa.bam -out sim.20.hsa.primary.bam `
`bedtools intersect -s -wo -bed -abam sim.20.hsa.primary.bam -b hsa.gff3 > sim.20.hsa.primary.bam.anno`
`python parse_bam.py`

# Results 

## filtering

Only 7283 mapped to the genome and were kept after filtering. And 3318
overlapped with any miRNA.

### detection

```{r iso}
library(ggplot2)
library(reshape)
data_gg <- melt(data_filtered,id.vars = c("seq","known","correct","mapped","amb"))
ggplot(data_gg, aes(correct, fill = value)) +
    geom_bar() +
    scale_fill_brewer(palette = "Set1") +
    theme_bw() +
    facet_wrap(~variable)
```


### multi-mapped

Many people filter out the sequences that map multiple times on the genome.
I count the times the sequence was mapping to another miRNA.

```{r m}
ggplot(data_filtered, aes(correct, fill = factor(amb))) +
    geom_bar() +
    scale_fill_brewer("number of times annotated",palette = "Set1") +
    theme_bw() 
```

## no filtering

16377 mapped to the genome. And 12549
overlapped with any miRNA.

### detection

```{r iso-no-filtered}
library(ggplot2)
library(reshape)
data_gg <- melt(data,id.vars = c("seq","known","correct","mapped","amb"))
ggplot(data_gg, aes(correct, fill = value)) +
    geom_bar() +
    scale_fill_brewer(palette = "Set1") +
    theme_bw() +
    facet_wrap(~variable)
```


### multi-mapped

Many people filter out the sequences that map multiple times on the genome.
I count the times the sequence was mapping to another miRNA.

```{r m-no-filtered }
ggplot(data, aes(correct, fill = factor(amb))) +
    geom_bar() +
    scale_fill_brewer("number of times annotated",palette = "Set1") +
    theme_bw() 
```
