---
title: "Get colors for your heatmap annotation"
output: 
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This post will show how to configure quickly the colors for the annotation
of rows/columns that go on top or on the side of a heatmap.

I normally use `pheatmap` a lot. Recently I discovered `ComplexHeatmap`. In both cases I spend always sometime changing the colors of the annotations. I ended up coding a function inside my package `DEGreport` to do that.

Given a data.frame with metadata data information will do:

* white-black scale for numerical variables
* blue-orange for categorical values with only two categories
* `Set2` palette for categorical values with more than two categories 

All three color sets can be changed using the [parameters function](http://lpantano.github.io/DEGreport/reference/degColors.html).

```{r}
library(DEGreport)
library(stemHypoxia)
data(stemHypoxia)
```

I am going to use this [stemcell experiment dataset](https://bioconductor.org/packages/release/data/experiment/html/stemHypoxia.html) I found randomly in Bioconductor page.

```{r}
exp = as.matrix(M[,3:30])
rownames(exp) = M[,2]
rownames(design) = colnames(exp)
design = design[,c(1:2)]
design$time = as.factor(design$time)
```

I'll find the most variable genes.

```{r}
most_variable = names(tail(sort(apply(exp, 1, sd)), 1000))
design$mean_gene = colMeans(exp[most_variable,])
```

And we can plot the heatmap with new colors with need to specify each variable at a time.

```{r, fig.height=8, fig.width=6}
library(pheatmap)
pheatmap(exp[most_variable,], scale = "row", show_rownames = FALSE,
         annotation_col = design,
         annotation_colors = degColors(design))
```

If you use `ComplexHeatmap` you can activate `col_fun = TRUE` to get it working.

* R Session

```{r}
devtools::session_info()
```


